### 2020-04-02 hamrony_select.Rmd
rank harmonies and ppt diagram update

---
title: "Ranking harmonies"
author: "Sayani Gupta"
date: "20/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gravitas)
library(philentropy)
```

# Ranking harmonies based on maximum paiwise distance between distributions of levels

Step 1:

For each harmony pair ($C_1, C_2$) with $m$ and $n$ levels respectively , assign a tibble with $C_1$ denoting rows and $C_2$ denoting columns. In the raw data structure, each cell of this $m*n$ order tibble would be a vector of multiple observations.

Step 2:

Since the length of each of these vectors inside each cell is different, the estimated density values of this data is obtained for n = 512(default?) for each cell. We still have a  $m*n$ order tibble with length of vector in each cell equal to 512 now.

Step 3:

For each of these $m$ levels of $C_1$, we need to compute a matrix which provides the pairwise distances of all the $n$ levels of $C_2$. Each of these matrix would be $n*n$ (if symmetric distance obtained just a traingular matric needs to be computed. Thus, in step 3 we have a list of $m$ elements, each of which is a matrix of order $n*n$.

Step 4:

For each of the  $m$ elements, maximum pairwise distance is obtained for each column of the $n*n$ matrix.  Thus, in step 4, maximum pairwise distance for each of the levels of $C_2$ are recorded. Again we have a (m*n) data frame with each cell as a scalar representing the maximum paiwise distances for each level of $C_2$.

Step 5:

Take maximum value of this matrix or take sum of the elements of this matrix to obtain a measure of variation for this harmony pair. The harmony pair with the maximum variation amongst these will be most interesting.

For example, let us take the harmony pair (wknd_wday, hour_day):

```{r}
library(tidyverse)
library(gravitas)
```

# all harmony pairs
```{r}
harmony_data <- smart_meter10 %>% harmony(
  ugran = "month",
  filter_out = c("hhour", "fortnight")
) 
```

# if they have same pairs but one placed on facet and other on x-axis
# then computational load can be minimised by

```{r create_gran}
 data_new <- smart_meter10 %>% 
   create_gran(harmony_data$facet_variable[7]) %>%
   create_gran(harmony_data$x_variable[7]) 
```

```{r step1}
step1 <- data_new[-c(1:2)]  %>% 
  pivot_wider(names_from = harmony_data$x_variable[1],
              values_from = general_supply_kwh,
              values_fn = list(general_supply_kwh = list))

step1
```

```{r step2}
density_extractx <- function(x)
{
  density(x)$y
}

colNms <- colnames(step1)[2:8]

step2 <- NULL

for (i in 1:length(colNms)) {
  step2[[i]] <- lapply(step1[[colNms[i]]], density_extractx)
}
str(step2)
```

```{r step3-4}
compute_JSD <- function(x, y)
{
  mat <- rbind(x, y)
  return(JSD(mat))
}

rowTibb <- step1[, 1]
step3 <- rep(list(diag(nrow(rowTibb))), length(colNms))
#step4 <- matrix(NA, ncol = nrow(rowTibb), nrow = length(colNms))
step4 <- array(NA, dim = nrow(rowTibb))
## Logic
# for each of the list 7 DOW
#__ find the stepped sum difference of density vector elements
for (k in 1:length(colNms)){

  dist <- matrix(NA, 
                 nrow = nrow(rowTibb),
                 ncol = nrow(rowTibb)) ## Matrix
  row_of_col_max <- NULL
  for(i in 1:nrow(rowTibb))
  {
    for (j in 1:nrow(rowTibb))
    {
      m1 <- step2[[k]][[i]]
      m2 <- step2[[k]][[j]]
      message(paste0("K:",k," I:",i," J:", j))
      dist[i, j] <- compute_JSD(m1, m2)
      # row_of_col_max[j] <- max(dist[, j]) 
       row_of_col_max <- max(dist)  
      # maximum of the entire matrix
    }
  }
  step3[[k]] <- dist
  step4[k] <- row_of_col_max
  step5 <- mean(step4)
}

step3
step4
step5
```