### 2020-05-14 algo_validation.Rmd
- compared norming constants for similar levels
- checked with other papers regarding norming constants

---
title: "Untitled"
author: "Sayani Gupta"
date: "14/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(gravitas)
library(tidyverse)
```

```{r}
cricket_data <- read_rds("data/cricket_data.rds")
cricket_data %>% hist_distance(gran1 = "lag_field", gran2 = "over", response = "runs_per_over", hierarchy_tbl = hierarchy_model)
```

```{r}

  hierarchy_model <- tibble::tibble(
      units = c("index", "over", "inning", "match"),
      convert_fct = c(1, 20, 2, 1))

.data <- cricket_data

hierarchy_tbl <- hierarchy_model

response <- "runs_per_over"

harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model)

smart_cric_normal <- cricket_data %>% 
  rank_harmony(harmony_tbl = harmonies_cric, 
               response = response,
               hierarchy_tbl = hierarchy_model,
               dist_distribution = "normal")

smart_cric_gamma <- cricket_data %>% 
  rank_harmony(harmony_tbl = harmonies_cric, 
               response = response,
               hierarchy_tbl = hierarchy_model,
               dist_distribution = "gamma")

smart_cric_weibull <- cricket_data %>% 
  rank_harmony(harmony_tbl = harmonies_cric, 
               response = response,
               hierarchy_tbl = hierarchy_model,
               dist_distribution = "weibull")
```

Now the pair (over, lag_field) and (over_inning_match) has the same levels (20,2). Hence we can see if the norming constants are working properly using these as example.

```{r}
cricket_data %>% 
  prob_plot("over",
            "lag_field",
            plot_type = "boxplot",
            response = response,
            hierarchy_tbl = hierarchy_model)
```

```{r}
cricket_data %>% 
  prob_plot("over",
            "inning_match",
            plot_type = "boxplot",
            response = response,
            hierarchy_tbl = hierarchy_model)
```

```{r}
cricket_data %>% 
  prob_plot("lag_field",
            "over_match",
            plot_type = "boxplot",
            response = response,
            hierarchy_tbl = hierarchy_model)
```

```{r}
cricket_data %>% 
  prob_plot("inning_match",
            "over_inning",
            plot_type = "boxplot",
            response = response,
            hierarchy_tbl = hierarchy_model)
```