### 2020-09-23 cover-letter/cover-letter.Rmd
minor changes

---
from: Sayani Gupta
fromqualifications: M.Stat
fromtitle: PhD student
fromWWW: sayani.netlify.app
fromemail: Sayani.Gupta@monash.edu
fromphone: +61 4 2388 9998
fromsignature: signature.png
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick,
closing: Best wishes
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider my article entitled “Visualizing probability distributions across bivariate cyclic temporal granularities” co-authored by Rob J Hyndman, Dianne Cook, and Antony Unwin for publication in the Journal of Computational and Graphical Statistics.

We believe that this manuscript is appropriate for publication by the Journal of Computational and Graphical Statistics because it makes unique computational and graphical contributions for analyzing temporal data and also non-temporal data with a nested hierarchical structure.

The manuscript is original. A preprint has been made available at arXiv and Monash EBS working paper series, but no part has been published before, nor is any part of it under consideration for publication at another journal.

Thank you for your consideration!